Title Goes Here

### 2021-04-09 analysis/paper/shceof.Rmd
Cambia a ERSST v5
Closes #28

We used monthly geopotential height, air temperature, ozone mixing ratio, and Total Ozone Column (TOC) at 2.5\degree\ longitude by 2.5\degree\ latitude of horizontal resolution and 37 vertical isobaric levels from ERA5 [@era5] for the period 1979 to 2019. We restricted most of our analysis to the post-satellite era to avoid any confounding factors arising from the incorporation of satellite observations, but we also used the preliminary back extension of ERA5 from 1950 to 1978 [@era5be] to look at long-term trends. We derived streamfunction at 200 hPa from ERA5 vorticity using the FORTRAN subroutine FISHPACK [@fishpack] and computed horizontal wave activity fluxes following @plumb1985. For precipitation data we used monthly data from the CPC Merged Analysis of Precipitation [@cmap], with a 2.5\degree\ resolution in latitude and longitude. This rainfall gridded dataset is based on information from different sources such as rain gauge observations, satellite inferred estimations and the NCEP-NCAR reanalysis, and it is available since 1979 to present.

@@

```{r sst}
sst <- ERA5_SST() %>%
  ReadNetCDF(vars = c("t" = "sst")) %>%
  normalise_coords() %>%
  .[season(time) == "SON"] %>%
  .[, w := season_weights(time)] %>% 
  .[, .(t = mean(t*w)), by = .(lon, lat, time = seasonally(time))]
```

```{r regr_sst}
regr_sst <- ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), 
    by = .(season, cEOF)] %>%
  .[sst, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>%
  .[, FitLm(t, Imaginary, Real, se = TRUE), by = .(lon, lat, cEOF, season)] %>%
  rm_intercept() %>%
  .[, term := factor_ReIm(term)]
```

@@

```{r sst-psi-2,  fig.cap = "(ref:sst-psi-2-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env="figure*"}
plot_sst_psi(which_cEOF = "cEOF2", which_lev = 200)
```

@@

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(which_cEOF = "cEOF1", which_lev = 200)
```

>>>>

We used monthly geopotential height, air temperature, ozone mixing ratio, and Total Ozone Column (TOC) at 2.5\degree\ longitude by 2.5\degree\ latitude of horizontal resolution and 37 vertical isobaric levels from ERA5 [@era5] for the period 1979 to 2019. We restricted most of our analysis to the post-satellite era to avoid any confounding factors arising from the incorporation of satellite observations, but we also used the preliminary back extension of ERA5 from 1950 to 1978 [@era5be] to look at long-term trends. We derived streamfunction at 200 hPa from ERA5 vorticity using the FORTRAN subroutine FISHPACK [@fishpack] and computed horizontal wave activity fluxes following @plumb1985. Sea Surface Temperature (SST) fields were taken from Extended Reconstructed Sea Surface Temperature (ERSST) v5 [@huang2017]. For precipitation data we used monthly data from the CPC Merged Analysis of Precipitation [@cmap], with a 2.5\degree\ resolution in latitude and longitude. This rainfall gridded dataset is based on information from different sources such as rain gauge observations, satellite inferred estimations and the NCEP-NCAR reanalysis, and it is available since 1979 to present.

@@

```{r sst}
sst <- ERSST() %>%
  readRDS() %>% 
  .[season(time) == "SON"] %>%
  normalise_coords() %>%
  .[, w := season_weights(time)] %>% 
  .[, .(t = mean(t*w)), by = .(lon, lat, time = seasonally(time))]
```

```{r regr_sst}
regr_sst <- ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), 
    by = .(season, cEOF)] %>%
  .[sst, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>%
  .[, FitLm(t, Imaginary, Real, se = TRUE), by = .(lon, lat, cEOF, season)] %>%
  rm_intercept() %>%
  .[, term := factor_ReIm(term)]

@@

```{r sst-psi-2,  fig.cap = "(ref:sst-psi-2-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env="figure*"}
plot_sst_psi(which_cEOF = "cEOF2", which_lev = 200)

@@

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(which_cEOF = "cEOF1", which_lev = 200)
